---
title: "activity"
author: "Prasad Sripragada"
date: "January 31, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Activity monitoring measures 

-Data is collected through electronic gadgets
-Data represents the steps taken on a day for every 5 min interval
-Data is storedin activy.csv file

###Loading the data
```{r echo=FALSE}
setwd("C:/coursera/datascience/Activity_monitoring_data")

activity <- read.csv("activity.csv")
```

### Check the data summary
```{r echo=TRUE}
summary(activity)
```

### Histogram of steps take each day
- Aggregate the sum of the steps taken on each day
- We are using sqldf library for this operation
- Plotting the histogram on the aggregated dataset

```{r echo=FALSE}
library(sqldf)
```

```{r echo=FALSE}

activity_per_day<-sqldf("select date,sum(steps) as TotalSteps from activity group by date")
activity_per_day$date<-as.Date(activity_per_day$date)


```

```{r echo=TRUE}
hist(activity_per_day$TotalSteps, breaks = 5,main="Steps taken each day", col="brown")
```


###Mean and Median of the steps taken 
```{r echo=TRUE}
steps_mean<-mean(activity$steps, na.rm = TRUE)
steps_median <- median(activity$steps, na.rm = TRUE)


```

Mean is `r steps_mean` and median is `r steps_median`

###Time series plot of average no of steps taken
- We use sqldf library to aggregate data
- Data is aggregate on interval and taking avg no of steps taken in those intervals

```{r echo=TRUE}
activity_mean_per_interval<-sqldf("select interval, avg(steps) AvgSteps from activity group by interval")
plot(activity_mean_per_interval$interval,activity_mean_per_interval$AvgSteps, type="l", main="Avg steps per interval")

```

###check which interval contains the max no of steps
```{r echo=FALSE}
max_interval <-sqldf("select interval from activity_mean_per_interval 
                     where AvgSteps=(select max(AvgSteps) from activity_mean_per_interval)")

```

The interval which contains the max no of steps is : `max_interval`


###Imputing the missing values in the data - the strategy applied
- Imputing on missing values is done on taking the avg steps from the intervals
- Impute the avg steps value from the non missing data average aggregated on intervals
- Seperate the missing value data and combine with avg values
- Once values are imputed combine with the non missing values data

```{r echo=FALSE}
activity_na <-sqldf("select * from activity where steps is null")
activity_not_na <-sqldf("select * from activity where steps is not null")

```
- On above code we seperated missing value data and non missing value data
- In following steps we make a new set by joining missing value data with average steps by interval data

```{r echo=FALSE}
activity_imputed <-sqldf("select b.AvgSteps as steps,a.date,a.interval from activity_na as a 
inner join activity_mean_per_interval as b on a.interval=b.interval")

```

- Now we combine the imputed data with missing values data
```{r echo=FALSE}
activity_combined<- rbind(activity_imputed,activity_not_na)
```

###Plot a histogram after imputing date- no of steps on each day
```{r echo=TRUE}
activity_combined_per_day<-sqldf("select date,sum(steps) as TotalSteps from activity_combined group by date")
activity_combined_per_day$date<-as.Date(activity_combined_per_day$date)
hist(activity_combined_per_day$TotalSteps, breaks = 5,main="Steps taken each day", col="brown")

```

###plot time series graph on imputed data by weekdays and weekends
- First make a new variable to name the day of the week on data variable
```{r echo=FALSE}
activity_combined$date<-as.Date(activity_combined$date)

activity_combined$weekday<-weekdays(activity_combined$date)

```

- Seperate the data sets based on weekdays and weekends
```{r echo=FALSE}
activity_weekday<-sqldf("select interval, avg(steps) AvgSteps from activity_combined where weekday not in ('Saturday','Sunday') group by interval")
activity_weekend<-sqldf("select interval, avg(steps) AvgSteps from activity_combined where weekday  in ('Saturday','Sunday') group by interval")

```

- Plot a graph
```{r echo=TRUE}
par(mfrow=c(2,1))

plot(activity_weekday$interval,activity_weekday$AvgSteps, type="l", main="Weekday avg steps in interval")
plot(activity_weekend$interval,activity_weekday$AvgSteps, type="l", main="Weekend avg steps in interval")

```

###Has the mean and median changed after imputing
```{r echo=TRUE}
steps_mean_imputed<-mean(activity_combined$steps)
steps_median_imputed <- median(activity_combined$steps)

change_mean<-steps_mean - steps_mean_imputed
change_median<-steps_median - steps_median_imputed

```

change in mean is `r change_mean` and change in median is ` r change_median`

